---
title: "Social Survey"
author: "Put your name here!"
date: "put the date here"
output:
  html_document: default
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
source("Import_GSS_2014_haven.R")
source('D:/Project_folders/GenSocSurvey/makeTables.R')
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->   

####This study deals with questions about possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data is from the 2014 general social survey.

The General Social Survey (GSS) is a survey that has been conducted every two years since 1972. The GSS monitors social change in American society and is one of the most important studies in the social sciencdes. The data collected in this particular study is based on the opinions of the resopndents interviewed in 2010 and were aslo later interviewed in 2010 and 2014. But, for the purpose of this report, we will be using only the data from 2014.

Our research question is: Was there a relationship bewtween gun ownership and race, sex, political views, opinions on capitol punishment, or belief in God in 2014?

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The data for the General Social Survey is large, having 2044 observations, 3542 variables, and includes not only the 2010 data, but data from 2012 and 2014. To make the data easier to use, we mut trim it down. We will be using r code stored in a separate .r file to help do this. We start with GSS data from the three years and then use the select function from the dplyr package to extract 12 variables that we think are important, but for this research papaer, we will only be using the following variables; belief in the gun ownership, race, sex, political beliefs, opinions on capitol punishment and belief in God. But, we still need to clean the dat up a bit more. This was done in the following ways:

  Gun ownership was originally recorded as YES, NO, IAP, REFUSED, and DK. Using the revalue function, we changed the IAP, REFUSED and DK value to NA.
  
  Race was originally recorded as having WHITE, BLACK, OTHER and IAP. Using the revalue function, we changed the IAP value to that of NA.
  
  For Sex, it originally had three values; MALE, FEMALE, and IAP. Again, using the revalue function, we changed the IAP to an NA value.
  
  For political beliefs, it was orginally recorded as having 6 different levels. Using the revalue function, we condensed the number of levels down to three. For the responses of EXTREMELY LIBERAL AND SLIGHTLY LIBERAL we changed the value to  LIBERAL. For the responses of SLIGHTLY CONSERVATIVE and EXTREMEMLY CONSERVATIVE we changed the value to CONSERVATIVE.
  
  Captial punishment was orignally recorded as IAP and NA. Using the revalue function, we changted the IAP value to that of NA.
  
  One's belief inGod was originally recorded as having 8 different levels. Again, using the revalue function, we condensed the number of leves down to four. For the responses DON'T BELIEVE and NO WAY TO FIND OUT, we changed the value to no. for the responses SOME HIGHER POWER, BELIEVE SOMETIMES, BELIEVE BUT DOUBTS, and DK (for don't know), we had them represented by the value maybe. For the response KNOW GOD EXISTS, we used the value yes. For the response IAP, we used the value NA.
  
To help answer our research question, we will be omitting all NA responses, and then view the results of the data both graphically and numercially. Since we will be comparing factor variables, we will use barcharts and twe-way talbes. Then, finally, we will do an inferential test using the chi-square test.


#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

```{r}
race_table=xtabs(~race + owngun, data=n2014,exclude = "NA")
barchartGC(race_table,type="percent", main="Race vs. Gun Ownership")
```
In viewing the data on the graph, it appears that WHITE, BLACk, and OTHER all responded with a vote of "no" to gun ownership. This means that people should not own guns themselves.


```{r}
sex_table=xtabs(~sex + owngun, data=n2014,exclude = "NA")
barchartGC(sex_table,type="percent", main="Sex vs.Gun Ownership")
```

In this graph in comparison indicated that both males and females say "no" to gun ownership.


```{r}
polviews_table=xtabs(~polviews + owngun, data=n2014,exclude = "NA")
barchartGC(polviews_table,type="percent", main= "Political Beliefs")
```

In comparison between these different groups we are seeing a lot of "no" between the groups.

```{r}
cappun_table=xtabs(~cappun + owngun, data=n2014,exclude = "NA")
barchartGC(cappun_table,type="percent")
```

consequently in this dataset, the idea of no ownership of guns stands strong, especially with those that believe in capital punishment.


```{r}
god_table=xtabs(~god + owngun, data=n2014,exclude = "NA")
barchartGC(god_table,type="percent", main= "Belief in God vs. Gun Ownership")
```

In relations between these groups, they are all saying "yes or no."

Based on the graph, the relationships shown between each group forms a survey that a lot of the participants are saying "no" to gun ownership. There is hardly any chance that one of these groups will change their response to yes. The only relationship is the comparison of belief in God and gun ownership. Each group had over 60% vote "no" and about 35% vote "yes".

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

To view the results in numerical data, we will use a two-way with row percents using the rowPerc function from the tigerstats package.

```{r}
rowPerc(race_table)
```

Those in the WHITE group, 30.01% voted YES on gun ownership, 60.99% voted NO.
Those in the BLACK group, 16.87% voted YES on gun ownership, 83.13% Voted NO.
Those in the OTHER group, 13.93% voted YES on gun ownership, 86.07% voted No.

Aftering observing the data numercially, both BLACK and OTHER strong voted NO on gun ownership, whereas the group WHITE had more YES on gun ownership compared to the two other groups.

```{r}
rowPerc(sex_table)
```

Those in the MALE group, 39.76% voted YES, 60.24% voted No, however, those in the FEMALE group,27.32% voted YES, 72.68 voted NO. We can see that more FEMALES voted "NO" on gun ownership than the MALES.

```{r}
rowPerc(polviews_table)
```


Of those in the LIBERAL group, 24.52% voted YES, 75.48% voted NO. Of those in the MODERATE group, 33.20% voted YES, 66.80% voted NO. Of those in the CONSERVATIVE group, 42.17% voted YES, 57.83% voted NO. Regardless of political views, all three groups want nothing to do with gun ownership.

```{r}
rowPerc(cappun_table)
```

Of those in the FAVOR group 41.44% voted YES, 58.56% voted NO. Of those in the OPPOSE group 20.58% voted YES, 79.42% voted NO. Although both groups have the majority decision of NO, those that oppose capital punishment want nothing to do with gun ownership. This could mean that if someone doesn’t believe in capital punishment, they obviously do not like guns either.

```{r}
rowPerc(god_table)
```

Comparing the No’s, MAYBE’s, and YES’s we can seee that about 30% voted YES on gun ownership and about 68% voted NO. We can see that there are more people who do not beleive in GOD that want nothing to do with gun ownership.

# Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


But, looking at the graphical and numerical results alone is not enough to tell us if these results also apply to the population at large. To do this, we will need to run an inferential test. The one we will use is the chi-square test.

We will now need to run four chi-square tests using the chisqtestGC function from the tigerstats package. We will run each test with the assumption that there is no relationship between belief in the Big Bang and the other variable in question; race, sex, political views, belief on capital punishment and belief in God.

Belief in Gun Ownership vs. Sex: We will form our two hypothesis

Belief in gun ownership is independent of Sex.
Belief in gun ownership and sex are related.(not independent).

```{r}
chisqtestGC(sex_table)
```

Our P-Value is 0. The chances of getting such a result are infinitely small if the null hypothes is true. We therefore choose to reject the null hypothesis. So it appears that Sex and belief in gun ownership are related.

Belief in Gun Ownership vs. Race Our two hypothesis:  Belief in gun ownership is independent of Race. Belief in gun ownership and Race are related. (not independent).

```{r}
chisqtestGC(race_table)
```

Our P-Value is 0. The chances of getting such a result are infinitely small if the null hypothesis is true. We therefore choose to reject the null hypothesis. So it appears that Race and belief in gun ownership are related.

Belief in Gun Ownership vs. Political views:
Our two hypothesis are: Belief in gun ownership is independent of political view. Belife in gun ownership and political views are related (not independent).

```{r}
chisqtestGC(polviews_table)
```

Our P-Value is 0. The chances of getting such a result are infinitesimally small if the null hypothesis is true. We therefore choose to reject the null hypothesis. So it appears that political views and belief in gun ownership are related.

Belief in Gun Ownership vs Belief in Capital Punishment: Our two hypothesis are:

Belief in gun ownership is independent of belief in capital punishment.

Belief in gun ownership and belief in capital punishment are related. (not independent).

```{r}
chisqtestGC(cappun_table)
```

Our P-Value is 0. The chances of getting such a result are infinitesimally small if the null hypothesis is true. We therefore choose to reject the null hypothesis. So it appears that belief in capital punishment and belief in gun ownership are related.

Belief in Gun Ownership vs. Belief in God: Our two hypothesis are:

Belief in gun ownership is independent of belief in God.
Belief in gun ownership and belief in God are related. (not independent).

```{r}
chisqtestGC(god_table)
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
